### Filtering

The symbol `|` works as an "or" operator, meaning it will return items that satisfy either one or both of the conditions before and after the operator.

```{r}
filter(flights, dest == "IAH" | dest == "HOU")
```

The `&` is an "and" operator, meaning both conditions need to hold. This is useful in conjunction with the `|`-operator, since otherwise you can simply just add filters after each other like this:

```{r}
filter(!is.na(flight2), !is.na(flight1))
```

Another operator `%in%` returns `TRUE` if an item exists inside a vector:

```{r}
flights %>% 
  filter(dest %in% c("IAH", "HOU")
```

To create the opposite logic, i.e. "not in c()", you can use `!`:

```{r}
flights %>%
  filter(!(dest %in% c("IAH", "HOU")))
```

### Selecting

Some basic selection functions from `dplyr` are `starts_with`, `ends_with` and `contains`.

```{r}
select(flights, starts_with("dep_"), starts_with("arr_"))
```

#### Regex (regular expressions)

The function `matches` uses *regular expressions*. These can be used to match precise patterns.

```{r}
select(flights, matches("^(dep|arr)_(time|delay)$"))
```

Some common Regex meta characters are:

-   `.` Matches any character except newline.

-   `^` Matches the beginning of a string.

-   `$` Matches the end of a string.

-   `[]` Matches any one of the characters inside the brackets.

-   `|` Logical OR.

-   `*` Matches 0 or more repetitions of the preceding character.

-   `+` Matches 1 or more repetitions.

-   `?` Matches 0 or 1 repetition (optional match).

-   `{n,m}` Matches between n and m repetitions.
