
### Plotting

We use `ggplot2` as the standard package for plotting, and the main function is `ggplot`. We supply a data frame to the first argument and an aesthetic mapping to the second argument. We add *layers* of plotting components using the plus sign. A simple example:

```{r}
ggplot(df, aes(x = x_variable, y = y_variable, colour = grouping_variable)) +
  geom_point()
```

Many types of layers may contain other data sets via the `data` argument and/or updated aesthetic mappings via the `mapping` argument. Data and mappings are typically inherited from the layer above if not specified in a new layer. There are many types of functions for making further adjustments to labels, titles, axes and other properties. A more complete example may look like this:

```{r}
ggplot(df, aes(x = x_variable, y = y_variable)) +
  geom_point() +
  geom_point(mapping = aes(x = new_x_variable, y = new_y_variable, colour = grouping_var),
             data = new_df) +
  xlab("X label") +
  ylab("Y label") +
  ggtitle("Title of plot") +
  theme_minimal()
```
#### Statistics

Most geoms come in pairs with complementary statistics arguments that are almost always used in concert. These functions can be used to retrieve the data that is used to generate the plot. For example, for these geoms:

```{r}
geom_smooth()
geom_dotplot()
geom_point()
geom_bar()
```
We have these `stat` functions respectively:

```{r}
stat_smooth()
stat_bindot()
stat_qq()
stat_count()
```
The corresponding stat functions can be found by reading the documentation with for example `?geom_smooth`.
