### Reading data

Sometimes data contains the delimiter. In that case, use quote argument to escape:

```{r}
read_csv("x,y\n1,'a,b'", quote="'")
```

### Reading multiple files

Instead of reading in every single file on its own, you can list all the files in a directory,
read each of them into a list and combine them into a data frame:

```{r}
paths <- list.files(path = "your_files", pattern = "*.csv")

list_of_data <- map(paths, read_csv)

single_df <- list_rbind(list_of_data)
```

### Reading data from the files names

When information is given in the name of the file itself, we can extract it using
the set_names function:

```{r}
paths <- list.files(path = "your_files", pattern = "*.csv")

paths %>% 
  set_names(basename) %>% 
  map(read_csv)
  
```

### Parsing data

There are various parsing functions from `lubdridate` library, such as `ymd()`. These functions can be used to convert objects into time and date formats. 

#### Dates and time

Examples:

```{r}
d1 <- "January 1, 2010"
d2 <- "2015-Mar-07"
d3 <- "06-Jun-2017"
d4 <- c("August 19 (2015)", "July 1 (2015)")
d5 <- "12/30/14" # Dec 30, 2014
t1 <- "1705"
t2 <- "11:15:10.12 PM"
```

Can be parsed respectively:

```{r}
mdy(d1)     #> [1] "2010-01-01"
ymd(d2)     #> [1] "2015-03-07"
dmy(d3)     #> [1] "2017-06-06"
mdy(gsub("\\(", "", gsub("\\)", "", d4)))  #> [1] "2015-08-19" "2015-07-01"
mdy(d5)     #> [1] "2014-12-30"
hm(t1)      #> [1] "17:05:00"
hms(t2)     #> [1] "23:15:10.12"
```

#### Pivoting

The library `tidyr` provides various ways to pivot data. An example of a pivot is from

```{r}
preg <- tribble(
  ~pregnant, ~male, ~female,
  "yes", NA, 10,
  "no", 20, 12
)
```

To a longer data format:

```{r}
preg_tidy2 <- preg %>%
  pivot_longer(c(male, female), names_to = "sex", values_to = "count", values_drop_na = TRUE)
```

#### Splitting columns

You can split a column into multiple columns using `separate()` and `extract()` from `tidyr`:

```{r}
tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  separate(x, c("one", "two", "three"))
```

In the above case, the second item has four "values" `"d,e,f,g"`. In this case, the `g` is dropped. There are various arguments in the functions to determine how to deal with extra or missing items like this.

#### Basic find and replace

You can perform a basic find-and-replace operation on a vector like this:

```{r}
y[y == “find”] <- “replace”
```
